Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-03-14 PAPER.tex
First crack at revising the signal loss methods section to reflect the fact that we are now using P_r

\subsubsection{Signal Loss Methodology} 
\label{sec:siglossmethod}
In short, our method consists of injecting an EoR-like signal into the data and then measuring the amount recovered, with the difference being accounted as signal loss.  To capture the full statistical likelihood of signal loss, one requires a quick way to generate many realizations of simulated $21$ 
cm signal visibilities. Here we use the same method as in \citetalias{ali_et_al2015}, where mock Gaussian noise visibilities (mock EoR signals) 
are filtered in fringe-rate space to retain only ``sky-like" time-modes. This signal is then added to the data.\footnote{One 
specific change from \citetalias{ali_et_al2015} is that we add this simulated signal into the analysis pipeline before the final fringe-rate filter is 
applied to the data. Previously, the addition was done after that final fringe-rate filter step.  This change results in an increased 
estimate of signal loss, %(by a factor of $\sim$$10$), 
likely due to the use of the fringe-rate filter as a simulator. However, this pipeline difference, while significant, is not the dominant reason why signal loss was under-estimated in \citetalias{ali_et_al2015} (the dominant reason is explained in the main text in Section \ref{sec:siglossmethod}).}

@@

\begin{equation}
\textbf{r} = \textbf{x} + \textbf{e}.
\end{equation}

@@

\begin{eqnarray}
\label{eq:sigloss}
P_{\rm out,\alpha} &\equiv& \widehat{\textbf{p}}_{r,\alpha} \nonumber\\%-\widehat{\textbf{p}}_{x,\alpha} \nonumber \\
&=& \text{M}^{\alpha}_{r}\textbf{r}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{r},% - \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x},
\end{eqnarray}
where, for illustrative purposes and notational simplicity, we have written these equations with diagonal normalization matrices M, even though for our numerical results we choose a non-diagonal matrix $\mathbf{M}$ as in Equation \eqref{eq:phat}. $P_{\rm in}$ represents the uniform-weighted power spectrum of $\textbf{e}$, our simulated EoR signal, and $\widehat{\textbf{p}}_{r}$ is the power spectrum of the
data plus injection. \cc{explain why we evaluate $P_{out}$ at $p_{r}$}
%In the limit where no significant cross correlations between the data and the injected signals are found, this will recover the amount of injected power spectrum remaining after the covariance weighting operation (i.e. the amount of background EoR power, $P_{\rm eor}$, not removed by the foreground down-weighting procedure). In other words, we make the anzatz that

@@

Expanding out Equation \eqref{eq:sigloss}, $P_{\rm out}$ becomes:

@@

\begin{eqnarray}
\label{eq:crossterm}
P_{\rm out,\alpha} &= &  \text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} + \text{M}^{\alpha}_{b}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}
^{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& 2 \text{M}^{\alpha}_{c}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} %\nonumber \\
%&-& \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}
In order to investigate the effect of each of these terms on signal loss, all three components are plotted in Figure \ref{fig:sigloss_terms} for two cases: inverse covariance weighting ($\textbf{R}_{r} = \widehat{\textbf{C}}_{r}^{-1}$) and uniform-weighted ($\textbf{R}_{r} = \textbf{I}$). We will now examine the behavior of this equation in the regimes of very weak and very strong injected signal.

@@

The correlation between $\widehat{\textbf{C}}_r^{-1}\textbf{e}$ and $\widehat{\textbf{C}}_r^{-1}\textbf{x}$ means that we cannot compute signal loss using a signal-only 
simulation, which would yield greater values for $P_{\rm out}$ and thereby underestimate signal loss. This argument is also made in \citet{switzer_et_al2015} in regards to estimating loss associated with the foreground cleaning of Green Bank Telescope (GBT) data. Therefore, in our revised 
signal loss computation we use the full quantity for $P_{\rm out}$ as defined in Equation \eqref{eq:sigloss}, which subtracts the 
weighted power spectrum of the data from the weighted power spectrum of data plus EoR.

@@

@@

We now shift our attention towards computing signal loss for the fringe-rate filtered PAPER-64 dataset. While our methodology 
outlined below is independent of weighting scheme, here we demonstrate the computation using inverse covariance weighting 
($\textbf{R} = \widehat{\textbf{C}}^{-1}$), the weighting scheme used in \citetalias{ali_et_al2015} which, with an empirically estimated covariance, leads to substantial loss. With this weighting, our 
expressions for $P_{\rm in}$ and $P_{\rm out}$ become:

@@

\cc{need edits in this section} Phrased in the context of Bayes' rule, we wish to find the posterior probability distribution $p(P_{\rm in} | 
\phat_{x})$, which is the probability of $P_{\rm in}$ given the uncorrected power spectrum estimate $\phat_{x}$. Given our ansatz from Section \ref{sec:siglossmethod} that $P_{\rm out} \equiv \widehat{\textbf{p}}_{r}$ is a good estimate of $\p_{x}$, we can write

\begin{equation}
\label{eq:Bayes}
p(P_{\rm in} | P_{\rm out}) \propto \evalat[\Bigg]{\mathcal{L} (  P_{\rm out} | P_{\rm in})}{P_{\rm out}=\widehat{\textbf{p}}_{x}}
  p(P_{\rm in}) ,
\end{equation}

\noindent where $\mathcal{L} $ is the likelihood function defined 
as the distribution of data plus injection ($P_{\rm out}$) given the injection $P_{\rm in}$, evaluated at the data level $\phat_x$.  We construct this distribution  
by fixing $P_{\rm in}$, and simulating our analysis pipeline for many realizations of the injected EoR signal 
consistent with this power spectrum. The resulting distribution is normalized such that the sum is unity, and the 
whole process can then be repeated for a different value of $P_{\rm in}$.

@@

Going forward, we treat every $k$-value separately in order to determine an upper limit on EoR per $k$. For every injection level, we smooth the distribution of $P_{\rm out}$ values using kernel density estimators (\citealt{scott2008kernel}), normalizing each one. Stitching all of them together results in a 2-dimensional smoothed transfer function, shown in Figure \ref{fig:sigloss_transfercurve} for inverse covariance weighting (left) and uniform weighting (right). Although we only show figures for one $k$-value, we note that 
the shape of the transfer curve is nearly identical for all $k$'s. This transfer function is the likelihood function in Bayes' rule, namely $\mathcal{L} (  P_{\rm out} | P_{\rm in})$. We then simply take a horizontal cut across at the data value $\widehat{\textbf{p}}_{x}$ to yield a posterior distribution for the signal. We compute the $95\%$ confidence interval (an upper limit on EoR) from these distributions.

>>>>

\subsubsection{Signal Loss Methodology} 
\label{sec:siglossmethod}
In short, our method consists of injecting an EoR-like signal into the data and then measuring how much of this injected signal would be detectable given any attenuation of this signal by the (lossy) data analysis pipeline.  To capture the full statistical likelihood of signal loss, one requires a quick way to generate many realizations of simulated $21$ 
cm signal visibilities. Here we use the same method as in \citetalias{ali_et_al2015}, where mock Gaussian noise visibilities (mock EoR signals) 
are filtered in fringe-rate space to retain only ``sky-like" time-modes. This signal is then added to the data.\footnote{One 
specific change from \citetalias{ali_et_al2015} is that we add this simulated signal into the analysis pipeline before the final fringe-rate filter is 
applied to the data. Previously, the addition was done after that final fringe-rate filter step.  This change results in an increased 
estimate of signal loss, %(by a factor of $\sim$$10$), 
likely due to the use of the fringe-rate filter as a simulator. However, this pipeline difference, while significant, is not the dominant reason why signal loss was under-estimated in \citetalias{ali_et_al2015} (the dominant reason is explained in the main text in Section \ref{sec:siglossmethod}).}

@@

\begin{equation}
\label{eq:rxe}
\textbf{r} = \textbf{x} + \textbf{e}.
\end{equation}

@@

\begin{eqnarray}
\label{eq:sigloss}
P_{\rm out,\alpha} &\equiv& \widehat{\textbf{p}}_{r,\alpha} \nonumber\\%-\widehat{\textbf{p}}_{x,\alpha} \nonumber \\
&=& \text{M}^{\alpha}_{r}\textbf{r}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{r},% - \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x},
\end{eqnarray}
where, for illustrative purposes and notational simplicity, we have written these equations with diagonal normalization matrices M, even though for our numerical results we choose a non-diagonal matrix $\mathbf{M}$ as in Equation \eqref{eq:phat}.

The quantity $P_{\rm in}$, defined by Equation \eqref{eq:Pin}, is a uniformly weighted estimator of the power spectrum of $\mathbf{e}$. Since there is no noise contribution to $\mathbf{e}$, it can be considered the power spectrum of this particular realization of the EoR; alternatively, it can be viewed as the true power spectrum of the injected signal up to cosmic variance fluctuations. The role of $P_{\rm in}$ in our analysis is to serve as a reference for the power spectrum that would be measured if there were no signal loss or other systematics. This is then to be compared to $P_{\rm out}$, which approximates the (lossy) power spectrum estimate that is outputted by our analysis pipeline prior to any signal loss adjustments. In principle, one could compute $P_{\rm out}$ using end-to-end simulations of the instrument and data analysis pipeline. In practice, however, such simulations may not accurately reflect real-life systematics and foregrounds. To overcome this obstacle, one can make the assumption that since the EoR signal is expected to be small, the data vector $\mathbf{x}$ itself is our best model of these contaminants. Making this assumption, the injected EoR signal $\mathbf{e}$ takes on the role of the true EoR signal, and the sum of $\mathbf{x}$ and $\mathbf{e}$ (i.e., Equation \eqref{eq:rxe}) replaces $\mathbf{x}$ as our model of the measured data.
\cc{explain why we evaluate $P_{out}$ at $p_{r}$} \acl{Added! See how you like it}
%In the limit where no significant cross correlations between the data and the injected signals are found, this will recover the amount of injected power spectrum remaining after the covariance weighting operation (i.e. the amount of background EoR power, $P_{\rm eor}$, not removed by the foreground down-weighting procedure). In other words, we make the anzatz that

@@

Under this injection framework, we can begin to see explicitly why there can be large signal loss. Expanding out Equation \eqref{eq:sigloss}, $P_{\rm out}$ becomes:

@@

\begin{eqnarray}
\label{eq:crossterm}
P_{\rm out,\alpha} &= &  \text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} + \text{M}^{\alpha}_{b}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}
^{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& 2 \text{M}^{\alpha}_{c}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} %\nonumber \\
%&-& \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}
In order to investigate the effect of each of these terms on signal loss, all three components are plotted in Figure \ref{fig:sigloss_terms} for two cases: \acl{can we go through and add the word ``empirical"? The phrase ``inverse covariance weight" should be reserved for when one is weighting by the true inverse covariance.} inverse covariance weighting ($\textbf{R}_{r} = \widehat{\textbf{C}}_{r}^{-1}$) and uniform-weighted ($\textbf{R}_{r} = \textbf{I}$). We will now examine the behavior of this equation in the regimes of very weak and very strong injected signal.

@@

The correlation between $\widehat{\textbf{C}}_r^{-1}\textbf{e}$ and $\widehat{\textbf{C}}_r^{-1}\textbf{x}$ means that we cannot compute signal loss using a signal-only 
simulation, which would yield greater values for $P_{\rm out}$ and thereby underestimate signal loss. This argument is also made in \citet{switzer_et_al2015} in regards to estimating loss associated with the foreground cleaning of Green Bank Telescope (GBT) data. \acl{The following line is no longer true}Therefore, in our revised 
signal loss computation we use the full quantity for $P_{\rm out}$ as defined in Equation \eqref{eq:sigloss}, which subtracts the 
weighted power spectrum of the data from the weighted power spectrum of data plus EoR.

@@

{\bf In general:} Signal loss due to couplings between the data and its weightings arise because our unnormalized quadratic estimator from Equation \eqref{eq:qhat} ceases to be a quadratic quantity, and instead contains higher order powers of the data. However, the normalization matrix $\mathbf{M}$ is generally derived assuming that the unnormalized estimator is quadratic in the data. Thus, neglecting couplings between data and its weightings causes power spectrum estimates to be incorrectly normalized. As argued in this section, the resulting bias will tend to bias the results \emph{low}, which is simply another way of describing signal loss.

@@

We now shift our attention towards computing upper limits on the EoR signal for the fringe-rate filtered PAPER-64 dataset in a way that accounts for signal loss. While our methodology 
outlined below is independent of weighting scheme, here we demonstrate the computation using inverse covariance weighting 
($\textbf{R} = \widehat{\textbf{C}}^{-1}$), the weighting scheme used in \citetalias{ali_et_al2015} which, with an empirically estimated covariance, leads to substantial loss. With this weighting, our 
expressions for $P_{\rm in}$ and $P_{\rm out}$ become:

@@

\cc{need edits in this section} \acl{Done!} Phrased in the context of Bayes' rule, we wish to find the posterior probability distribution $p(P_{\rm in} | 
P_{\rm out})$, which is the probability of $P_{\rm in}$ given the uncorrected/measured power spectrum estimate $P_{\rm out}$. Bayes' rule gives

\begin{equation}
\label{eq:Bayes}
p(P_{\rm in} | P_{\rm out}) \propto {\mathcal{L} (  P_{\rm out} | P_{\rm in})}
  p(P_{\rm in}) ,
\end{equation}

\noindent where $\mathcal{L} $ is the likelihood function defined 
as the distribution of data plus injection ($P_{\rm out}$) given the injection $P_{\rm in}$.  We construct this distribution  
by fixing $P_{\rm in}$, and simulating our analysis pipeline for many realizations of the injected EoR signal 
consistent with this power spectrum. The resulting distribution is normalized such that the sum over $P_{\rm out}$ is unity, and the 
whole process is then repeated for a different value of $P_{\rm in}$.

@@

Going forward, we treat every $k$-value separately in order to determine an upper limit on EoR per $k$. For every injection level, we smooth the distribution of $P_{\rm out}$ values using kernel density estimators (\citealt{scott2008kernel}), normalizing each one. Stitching all of them together results in a 2-dimensional smoothed transfer function, shown in Figure \ref{fig:sigloss_transfercurve} for inverse covariance weighting (left) and uniform weighting (right). \acl{Why not use $P_{\rm out}$ on the vertical axes of these plots?} Although we only show figures for one $k$-value, we note that 
the shape of the transfer curve is nearly identical for all $k$'s. This transfer function is the likelihood function in Bayes' rule, namely $\mathcal{L} (  P_{\rm out} | P_{\rm in})$. After multiplying by the prior \acl{right?}, we then simply take a horizontal cut across at the data value $P_{\rm out}$ to yield a posterior distribution for the signal. We compute the $95\%$ confidence interval (an upper limit on EoR) from these distributions.
